---
title: "SIHF Game Detail Data"
author: "Markus Senn"
date: "7/9/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sihfapi)
library(httr)
library(tidyverse)
library(lubridate)
library(stringr)

# Converst first character of string to upper case (leaving all others as is)
firstToUpper <- function(x) {
  first <- toupper(substr(x, 1, 1))
  rest <- substr(x, 2, nchar(x))
  paste0(first, rest)
}
```

## The SIHF API

The website of the Swiss Ice Hocke Federation (SIHF) exposes an API for use by its "Gamecenter". The present package makes use of this API to download results and game details.

There does not seem to be any public documentation on the API and how to use it. Thus, the package has been developed through backwards engineering. This fact is still much noticable at some points and contributions are highly welcome.


## Fetch results

"Results" are short tables that summarise game outcomes. They are particularly important as we can retrieve long lists of results using handy filters such season, league and date.

```{r results}
results <- fetch_results(c(Season = 2017)) %>% 
  print()
```

Filters are not fully understood. The 'Date' filter has a somewhat particular behvior (see `?fetch_results()`).

```{r resultsDateFilter}
fetch_results(c(Season = 2017, Date = "20.9.2016"))
fetch_results(c(Season = 2017, Date = "1.10.2016-31.10.2016"))
```


Internally, the results object is a list of `c("content", "path", "query", "response")`.  

- content: parsed content returned by the API call
- path: path to the API endpoint querried ("/statistic/api/cms/table")
- query: API query elements; notably the filter criteria
- response: the raw http response of the API call (httr object)


## Results content

The content is provided by SIHF. It is a list object with multiple nested lists. At the top level, the list contains 14 elements:

```{r resultsContentTop}
str(results$content, max.level = 1)
```

Most of these elements are unintersting as they only contain rendering instructions for the SIHF website. By far the most interesting element `$data` which contains the actual data content of the response. `$header` provides meta data for each item of the data. This meta data is the name, description, and alias of each item as well as some item-specific rendering instructions.

```{r resultsContentHeader}
header <- tibble(
  name = map_chr(results$content$header, 'name'),
  description = map_chr(results$content$header, 'description'),
  alias = map_chr(results$content$header, 'alias') 
) %>% 
  print()
```

`$data` is a list of variable length, where each element of the list represents an individual game. Each of these games consists of 11 elements that correspond to the header items discussed above:

```{r resultsContentData}
d <- results$content$data %>% 
  map(~set_names(.x, header$alias))
  
d[[1]]
```

```{r resultsContentDataParsed1}
d <- d %>% 
  map(~enframe(.x) %>% 
        spread(name, value)
  ) %>% 
  bind_rows() %>% 
  mutate_at(vars(date, day, decision, status, time), unlist) %>% 

  # Parse date_time
  mutate(date_time = dmy_hm(paste(date, time))) %>% 
  select(-date, -day, -time) %>% 
  print()
```

This simple parsing contains nested lists and may be useful for further processing, as it contains a maximum amount of data with minimum structure. Further parsing may be more useful for human readers:

```{r resultsContentDataParsed2}
collapseScore <- function(score) {
  map2(
    score['homeTeam'], score['awayTeam'],
    ~paste(.x, .y, sep = ":")
  ) %>% 
    unlist(use.names = FALSE) %>% 
    paste(collapse = " | ")
}

results.games <- d %>% 
  mutate(
    gameId = map_chr(details, 'gameId'),
    awayTeamAcronym = map_chr(awayTeam, 'acronym'),
    homeTeamAcronym = map_chr(homeTeam, 'acronym'),
    finalScore = map_chr(score, collapseScore),
    periodScore = map_chr(scorePerPeriod, collapseScore)
  ) %>% 
  select(
    date_time, gameId,
    ends_with('TeamAcronym'),
    ends_with('Score', ignore.case = FALSE),
    decision
  ) %>% 
  print()
```

Note that the list of games we retrieved is neither resticted to NLA games nor a full list of all games available in the API. Rather, it appears to be a combination of NLA, CHL and national team games.

We can also use the results table to compile a simple teams table that lists teamId, acronym and name.

```{r resultsContentTeamsList}
results.teams <- d %>% 
  select(ends_with('Team')) %>% 
  gather("key", "value") %>% 
  select(value) %>% 
  mutate(value = map(
    value, ~enframe(.x) %>% 
      spread(name, value)
  )) %>% 
  unnest() %>% 
  mutate_all(unlist) %>% 
  select(id, acronym, name) %>% 
  distinct() %>% 
  arrange(id) %>% 
  print()
```

One of the important use case of the results table is to retrieve gameIds which we require to retrieve game details.


## Fetch game detail

Once we know the ID of a game we are interested in, we can easily fetch the detail of this game:

```{r gd_fetch}
gameDetail <- 
  results.games %>% 
  slice(30) %$%
  gameId %>% 
  fetch_gameDetail() %>% 
  print()
```

Internally, the results object is a list of `c("content", "path", "query", "response")`. These items are largely analogous to the ones in 'results'. Obviously, content is different.

### Multiple games

Using functions of the `apply` family or from the `purrr` package, fechting details on multiple games is simple.

```{r gd_multiFetch}
gameDetail <- 
  results.games %>% 
  slice(30:32) %$% 
  gameId %>% 
  map(fetch_gameDetail) %>% 
  map(print)
```

## Game detail content

The content of each game detail is a fairly large and complex list of lists:

```{r gd_content}
str(gameDetail[[1]]$content, max.level = 1)
```

Upon closer inspection, we find a number of items in this list that are relatively easy to transform into a table where each row represents a game and is identified by `gameId`. We divided these items into three classes that need separate parsing:

### Game

```{r gd_gameNames}
nm.game <- list(
  single = c("gameId", "season", "updateDateTime", "startDateTime", "endDateTime", "videoId", "info", "alias"),
  idName = c("league", "region", "group", "tournament", "qualification", "phase"),
  status = "status"
) %>% 
  print()
```

`single` is a group of items that consists only of a single value:

```{r gd_gameSingle}
gameDetail.game <- list()

gameDetail.game$single <- 
  gameDetail %>% 
  map('content') %>% 
  map(`[`, nm.game$single) %>% 
  map(enframe) %>% 
  map(~spread(., name, value)) %>% 
  bind_rows() %>% 
  mutate_all(unlist) %>% 
  mutate_at(
    vars(ends_with("DateTime")),
    ymd_hms
  ) %>% 
  print()
```

`idField` items consist of an integer ID and a matching name. For example `league = list(id = 81, name = "CHL")`:

```{r gd_gameIdField}
gameDetail.game$idName <- 
  gameDetail %>% 
  map('content') %>% 
  set_names(gameDetail.game$single$gameId) %>% 
  enframe("gameId", "data") %>% 
  mutate(
    data = data %>% 
      map(`[`, nm.game$idName) %>% 
      map(enframe)
  ) %>% 
  unnest() %>% 
  rename(var = name) %>% 
  mutate(
    value = value %>% 
      map(bind_cols)
  ) %>% 
  unnest() %>% 
  gather("key", "value", -gameId, -var) %>%
  mutate(key = paste0(var, firstToUpper(key))) %>% 
  select(-var) %>%
  spread(key, value) %>% 
  print()
```

That last item we will flatten to a table with one row per game is `status`. This item is somewhat more complex, consisting of 7 fields. We prefix these seven fields as "statusX".

```{r gd_gameStatus}
gameDetail.game$status <- 
  gameDetail %>% 
  map('content') %>% 
  set_names(gameDetail.game$single$gameId) %>% 
  enframe("gameId", "data") %>% 
  mutate(
    data = data %>% 
      map(`[`, "status") %>% 
      map(bind_cols)
  ) %>% 
  unnest() %>% 
    mutate_at(
    vars(ends_with("DateTime")),
    ymd_hms
  ) %>% 
  set_names(str_c("status", names(.) %>% firstToUpper())) %>% 
  print()
```

We join the three game items to a single data frame.

```{r gd_gameJoin}
gameDetail.game <- 
  gameDetail.game$single %>% 
  full_join(gameDetail.game$idName, by = "gameId") %>% 
  full_join(gameDetail.game$status, by = c(gameId = "statusGameId")) %>% 
  print()
```

### Complex
The remaining items of game detail content are more complex. Flattening them to a row-per-game table seems inappropriate. Doing so would substantially harm readability and handling of the data. We parse them into separate tables.

#### Summary

The `summary` object contains event data on each period (fouls, goals, goalkeepers) and the shootout. We parse these data items into separate tables to take accound of the fact that we have vastly different information on the differen events. 

Todo: The `goalkeepers` object is not yet parsed as it was empty for the games we have looked at so far.

```{r gd_summary}
# Parse one summary$periods object
parsePeriods <- function(periods) {
  ## Parse list
  tmp <- periods %>% 
    enframe() %>% 
    transmute(
      period = value %>% 
        map_chr("name"),
      
      fouls = value %>% 
        map("fouls") %>% 
        map(~map_df(., bind_cols)),
      
      goals = value %>% 
        map("goals") %>% 
        map(~map_df(., bind_cols)),
      
      goalkeepers = value %>% 
        map("goalkeepers")
    )
 
  ## Prepare flat return
  list(
    fouls = tmp %>% 
    select(period, fouls) %>% 
    unnest(),
    
    goals = tmp %>% 
    select(period, goals) %>% 
    unnest()
  ) 
}

tmp <- 
  gameDetail %>%
  map('content') %>% 
  set_names(gameDetail.game$gameId) %>% 
  enframe("gameId", "data") %>% 
  mutate(
    data = data %>%
      map("summary"),
    
    periods = data %>% 
      map("periods") %>% 
      map(parsePeriods),
    
    fouls = periods %>% 
      map("fouls"),
    
    goals = periods %>% 
      map("goals"),
    
    shootout = data %>% 
      map("shootout") %>% 
      map("shoots") %>% 
      map(~map_df(., bind_cols))
  ) %>% 
  select(-data, -periods)

gameDetail.fouls <- tmp %>% 
  select(gameId, fouls) %>% 
  unnest() %>% 
  print()

gameDetail.goals <- tmp %>% 
  select(gameId, goals) %>% 
  unnest() %>% 
  print()

gameDetail.shootout <- tmp %>% 
  select(gameId, shootout) %>% 
  unnest() %>% 
  print()

rm(tmp)
```
